---
title: "gamma_lrt_nonsim"
author: "Olai Gaarn Skogen"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(VGAM)
library(distr)
```

```{r}
?uniroot
```


We first take the case where n = 1, because then we have the distribution function for $\operatorname{log}(X)$ if $X \sim \text{Gamma}(\alpha, \beta)$.

We can go from 
1. $T(X) = \operatorname{log}(X)$ sufficient statistic to
2. $l(X; \alpha) = \alpha \sum_i{T(X_i)} - n \operatorname{log}(\alpha)$ likelihood to
3. $\hat{\alpha} = \frac{n}{\sum_i{T(X_i)}}$ mle to
4. $\Lambda(X) = \exp\{l(X; \hat{\alpha}) - l(X; \alpha_0) \}$ likelihood ratio

If we can find the critical value numerically, then we can also find the power numerically.

Pretty sure we did something similar in StatInf.

# Everything is for the Gamma function - test for shape parameter

```{r}
# find critical values for c
log_likelihood <- function(t, alpha){
  alpha * sum(t) - length(t) * digamma(alpha)
}

likelihood_ratio_calc <- function(t, alpha_0){
  alpha_hat = igamma(mean(t))
  log_lr =  log_likelihood(t, alpha_hat) - log_likelihood(t, alpha_0)
  log_lr
}

## finding t_dual, i.e. the value of t that gives the same LR value
find_t_dual <- function(t, alpha_0){
  # Assuming t_dual is higher than t
  # TODO : finding a way to make it go both ways
  lr <- likelihood_ratio_calc(t, alpha_0)
  
  modified_func <- function(t) {likelihood_ratio_calc(t, alpha_0) - lr}
  
  t_dual <- uniroot(modified_func, lower = t + 1e-4, upper = 10, extendInt = "no")$root
  t_dual
}

t = log(1)
log_likelihood(t, 2)
likelihood_ratio_calc(t, 2)

likelihood_ratio_calc(log(3), 2)

find_t_dual(t, 2)

## actually finding the critical value c

find_extremity_prob <- function(t, alpha_0){
  t_dual <- find_t_dual(t, alpha_0)
  1 - plgamma(t_dual, shape = alpha_0) + plgamma(t, shape = alpha_0)
}

find_extremity_prob(t, 2)

find_critcal_vals <- function(significance_level, alpha_0){
  modified_func <- function(t) {find_extremity_prob(t, alpha_0) - significance_level}
  
  t1 <- uniroot(modified_func, lower = -10, upper = log(1))$root
  t2 <- find_t_dual(t1, alpha_0)
  
  c1 = likelihood_ratio_calc(t1, alpha_0)
  c2 = likelihood_ratio_calc(t2, alpha_0)
  
  c(t1, t2, c1, c2)
}

crit_vals = find_critcal_vals(0.05, 2)
crit_vals
# find critical values t1 and t2


# find power
find_power <- function(alpha_true, t1, t2){
  1 - plgamma(t2, shape = alpha_true) + plgamma(t1, shape = alpha_true)
}

```

```{r}
start = 1.8
end = 2.2
n = 21

alpha_values <- seq(start, end, length.out = n)

pwr_values <- sapply(alpha_values, find_power, t1 = crit_vals[1], t2 = crit_vals[2])

plot(alpha_values, pwr_values)
```

```{r}

```




