---
title: "R Notebook"
output:
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    fonttheme: "structurebold"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(reshape2)
library(tidyverse)
```


## Model

```{r}
dmodel <- function(x, n, p){
  dbinom(x, n, p)
}

pmodel <- function(x, n, p, tail_left = TRUE){
  pbinom(x, n, p, tail_left)
}

sample_space <- function(n){
  0:n
}

expected_val_model <- function(n, p){
  n*p
}

```

## Methods for computing the p-value

```{r}
p_val_a <- function(x, n, p){
  pr_tail_l = pmodel(x, n, p, tail_left = TRUE)
  ## Use 1 - tail prob but add point prob
  pr_tail_r = 1 - pr_tail_l + dmodel(x, n, p)
  
  p_val = 2 * min(pr_tail_l, pr_tail_r, 1/2)
  p_val
}

p_val_a(7, 10, 0.7)
```

```{r}
p_val_b <- function(x, n, p, multi_modal = TRUE){
  W_x = dmodel(x, n, p)
  
  S = sample_space(n)
  p_val = 0
  
  if (multi_modal){
    # If multimodal makes this a bad for this p-method
    for(y in S){
      W_y = dmodel(y, n, p)
      if(W_y > W_x){
      }
      else{
        p_y = W_y
        p_val = p_val + p_y
      }
    }
  }
  else{
    #WIP of effectivised computation
  }
  
  p_val
}

p_val_b(3, 10, 0.3, TRUE)
```

```{r}
# Assuming point mass function
p_val_c <- function(x, n, p){
  p_tail_l = pmodel(x, n, p, tail_left = TRUE)
  p_tail_r = 1 - p_tail_l + dmodel(x, n, p)
  epsilon = 1e-8
  if (p_tail_l < p_tail_r + epsilon){
    # Add the maximum right tail that is smaller than left tail
    ## Start with 0 tail and add one point mass at a time until it grow bigger than left tail
    r_tail_max_smaller = 0
    y = n
    sugg = r_tail_max_smaller
    while(sugg < p_tail_l + epsilon){
      r_tail_max_smaller = sugg # Accept suggestion
      r_add = dmodel(y, n, p) # Calculate new point mass to be added
      sugg = sugg + r_add # Propose new suggestion
      y = y - 1 # Take one lower y-value
    }
    p_val = p_tail_l +  r_tail_max_smaller
  }
  else{
    # Do the same thing but find maximum smaller left tail instead of right
    l_tail_max_smaller = 0
    y = 0
    sugg = l_tail_max_smaller
    while(sugg < p_tail_r + epsilon){
      l_tail_max_smaller = sugg # Accept suggestion
      l_add = dmodel(y, n, p) # Calculate new point mass to be added
      sugg = sugg + l_add # Propose new suggestion
      y = y + 1 # Take one higher y-value
    }
    p_val = p_tail_r +  l_tail_max_smaller
  }
  if(p_val > 1){ # If the distrbution is symmetric it will count the central event twice.
    return(1)
  }
  p_val
}

p_val_c(7, 10, 0.68)
```

```{r}
p_val_d <- function(x, n, p) {
  exp_val = expected_val_model(n, p)
  W_x = abs(x - exp_val)
  
  p_val = 0
  S = sample_space(n)
  for(y in S){
    W_y = abs(y - exp_val)
    if(W_y < W_x){}
    else{
      p_val = p_val + dmodel(y, n, p)
    }
  }
  p_val
}

p_val_d(7, 10, 0.7)
```

```{r}
observator <- function(x, n, p){
  pr_tail_l = pmodel(x, n, p)
  pr_tail_r = 1 - pr_tail_l + dmodel(x, n, p)
  min(pr_tail_l, pr_tail_r)
}

p_val_obs <- function(x, n, p){
  W_x = observator(x, n, p)
  S = sample_space(n)
  
  p_val = 0
  epsilon = 1e-8
  
  for(i in S){
    W_i = observator(i, n, p)
    if(W_i < W_x + epsilon){
      p_val = p_val + dmodel(i, n, p)
    }
  }
  p_val
}

p_val_obs(3, 10, 0.3)
```


## Testing some different models

### Binomial with p = 0.5

```{r}
dmodel <- function(x, n, p){
  dbinom(x, n, p)
}

pmodel <- function(x, n, p, tail_left = TRUE){
  pbinom(x, n, p, tail_left)
}

sample_space <- function(n){
  0:n
}

expected_val_model <- function(n, p){
  n*p
}

```

```{r}
comparison_p_vals <- function(n, p){
  p_a = NULL
  p_b = NULL
  p_c = NULL
  p_d = NULL

  for(i in 0:n){
    p_a = c(p_a, p_val_a(i, n, p))
    p_b = c(p_b, p_val_b(i, n, p))
    p_c = c(p_c, p_val_c(i, n, p))
    p_d = c(p_d, p_val_d(i, n, p))
  }
  
  df <- data.frame(p_a, p_b, p_c, p_d)
  colnames(df) <- c("​Double tail", "Point", "Tail", "Distance")  #Added zero width space for "Double tail" to be shown first in plot
  df["x_val"] <- 0:n
  
  df
}

df <- comparison_p_vals(10, 0.5)

df %>%
  pivot_longer(-x_val, names_to = "Method", values_to = "P") %>%
  ggplot(aes(x = factor(x_val), y = P, fill = Method), xlab = "Method") +
  geom_bar(stat = "identity", width = .75, position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme_classic() +
  labs(x = "X", y = "p-value")
```


### Binomial with p = 0.7

```{r}
df <- comparison_p_vals(10, 0.7)

df %>%
  pivot_longer(-x_val, names_to = "Method", values_to = "P") %>%
  ggplot(aes(x = factor(x_val), y = P, fill = Method), xlab = "Method") +
  geom_bar(stat = "identity", width = .75, position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme_classic() +
  labs(x = "X", y = "p-value")
```
```{r}
df <- comparison_p_vals(10, 0.9)

df %>%
  pivot_longer(-x_val, names_to = "Method", values_to = "P") %>%
  ggplot(aes(x = factor(x_val), y = P, fill = Method), xlab = "Method") +
  geom_bar(stat = "identity", width = .75, position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme_classic() +
  labs(x = "X", y = "p-value")
```
### Binomial with n = 30, p = 0.7
```{r}
df <- comparison_p_vals(30, 0.7)

df %>%
  pivot_longer(-x_val, names_to = "Method", values_to = "P") %>%
  ggplot(aes(x = factor(x_val), y = P, fill = Method), xlab = "Method") +
  geom_bar(stat = "identity", width = .75, position = "dodge") +
  theme_classic() +
  labs(x = "X", y = "p-value")+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  ylim(0, 0.00005)
  
```

### Binomial with n = 30, p = 0.9
### Binomial with n = 50, p = 0.7
### Binomial with n = 50, p = 0.9
### Binomial with n = 100, p = 0.7
### Binomial with n = 100, p = 0.9


### Binomial with n = 40 p = 0.82

```{r}
n = 40
p = 0.82

df <-comparison_p_vals(n, p)

df %>%
  pivot_longer(-x_val, names_to = "Method", values_to = "P") %>%
  ggplot(aes(x = factor(x_val), y = P, fill = Method), xlab = "Method") +
  geom_bar(stat = "identity", width = .75, position = "dodge") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))+
  theme_classic() +
  labs(x = "X", y = "p-value")
  
  

# Plotting 
df_points = data.frame(x = 0:n, y = dbinom(0:n, n, p))
ggplot(df_points, aes(x = x, y = y)) +
  geom_bar(stat = "identity")
```


### Poisson with lam = 1

### Poisson with lam = 2

### Poisson with lam = 4

### Negative binomial 

### Hypergeometric


#### Unhyperness = 1, p = 0.5

```{r}
dmodel <- function(x, n, p, unhyperness = 1){
  k = n
  m = round(unhyperness * k)
  n_ = round(m * (1-p)/p)
  dhyper(x, m, n_, k)
}

pmodel <- function(x, n, p, unhyperness = 1, tail_left = TRUE){
  k = n
  m = round(unhyperness * k)
  n_ = round(m * (1-p)/p)
  val = phyper(x, m, n_, k, lower.tail = tail_left)
  
  if(is.nan(val)){
    0
  }
  else{
    val
  }
}

sample_space <- function(n){
  0:n
}

expected_val_model <- function(n, p){
  n*p
}

?floor
```

```{r}
k = 10
n_ = round(1 * k)
m = round(n_ * 0.5 / 0.5)

phyper(10, m, n_, k, FALSE)

pmodel(10, 10, 0.5, FALSE)
```

```{r}
observator <- function(x, n, p){
  pr_tail_l = pmodel(x, n, p)
  pr_tail_r = 1- pr_tail_l + dmodel(x, n, p)
  min(pr_tail_l, pr_tail_r)
}

p_val_a(7, 10, 0.5)
p_val_b(7, 10, 0.5)
p_val_c(7, 10, 0.5)
p_val_d(7, 10, 0.5)

comparison_p_vals(10, 0.47)
```


#### Unhyperness = 1, p = 0.7

```{r}
comparison_p_vals(10, 0.68)
```

```{r}
pmodel(3, 10, 0.7)
pmodel(3, 10, 0.7, tail_left = FALSE) + dmodel(3, 10, 0.7)
```


#### Unhyperness = 1, p = 0.9

```{r}
df <- comparison_p_vals(10, 0.9)

df %>%
  pivot_longer(-x_val, names_to = "variable", values_to = "P") %>%
  ggplot(aes(x = x_val, y = P, fill = variable), xlab = "Method") +
  geom_bar(stat = "identity", width = .75, position = "dodge")
```


### Negative hypergeometric

### Fishers exact test

```{r}
fisher_p_val_test <- function(n_x, n_y){
  c_vec = 0:(n_x + n_y)
  
  p_df = data.frame(matrix(NA, nrow = n_x + n_y, ncol = n_x + n_y))
  rownames(p_df) <- 0:(n_x + n_y)
  colnames(p_df) <- 0:(n_x + n_y)
  for(c in c_vec){
    x_vec = 0:c
    for(x in x_vec){
      #Calculate p value
      #Set p_value at place p_df[x, c]
    }
  }
}
```


## Computing power analytically

```{r, eval=FALSE}


power_computation <- function(n, p_h0, alpha, p_val_meth){
  S = sample_space(n)
  for (y in S){
    p_y = p_val_meth(y, n, p_h0)
    if 
  }
}
```

